---
title: "The landscape of biotechnology ETFs"
description: |
  Exploring stock pickers' view of value in biotechnology
author: Josh Peters
date: 02-27-2021
categories:
  - Industry
  - Finance
output:
  distill::distill_article:
    code_folding: true
    self_contained: false
    toc: true
draft: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(glue)
library(tidyquant)
library(ggiraph)
root <- "/Users/jpeters/Downloads"

```

## The Genomic Revolution  
Genomic Revolution (ARKG) is a high-performing ETF listed among the popular thematic ARK Invest ETFs. It has recently provided enormous returns, prompting large inflows. ARKG states portfolio companies focus on the following technology areas: DNA sequencing, gene editing, targeted therapeutics, bioinformatics, and agricultural biology.

After listening to ARK’s Head of Research, their approach is attractive, simple, and focused on the underlying companies’ 5-year potentials. His sentiment echoed much of the value investing ethos of Graham mixed with General Purpose Technology theory and a comforting humility. The idea here is that ARKG’s holdings will likely fall into two bins: unique golden tickets, or consensus picks. This assumes independence between funds and, while that probably isn’t true, I assume here that funds have their own autonomous reasons for choosing their portfolio. 

Given the hype around ARK Invest’s thematic ETFs, I wanted to visualize where these ETFs reside compared to other ETFs. To do so, I prepared a description of all ETFs’ holdings and performance (1-year alpha and beta). Data is downloaded from [masterdatareports.com](https://masterdatareports.com/).

```{r echo=FALSE}
# load dat
dat <- read_csv(file = glue("{root}/ConstituentData42 (1).csv"))
dat <- janitor::clean_names(dat)
dat <- dat[, c("sponsor", "composite_ticker", "composite_name", "constituent_ticker",
                 "constituent_name", "weighting", "date", "last_trade", "bloomberg_symbol",
                 "naics_sector", "naics_sub_industry", "market_cap", "earnings",
                 "pe_ratio", "dividend_amt", "dividend_yield")]
rmarkdown::paged_table(dat[1:100, ])

# load etf information
etfs <- read_csv(glue("{root}/AllTypes.csv"), col_names = FALSE)
etfs <- etfs[, -4]
colnames(etfs) <- c("name", "ticker", "index")

```

## ETF Embedding  
First, to visualize the compositional ETF data, we utilized UMAP to embed the composition of thematic ETFs, emphasized by restricting our list to funds with less than 60 holdings, in a two-dimensional representation.

```{r echo=FALSE}

dat <- dat[!is.na(dat$constituent_ticker), ]
dat$constituent_ticker <- gsub("[^\u0001-\u007F]+|<U\\+\\w+>","", dat$constituent_ticker)
dat <- dat[dat$constituent_ticker != "", ]
dat <- dat[dat$weighting != "parent", ]
dat <- dat[!grepl("CASH", dat$constituent_ticker), ]
dat <- dat[!grepl("Cash", dat$constituent_ticker), ]
dat$weighting <- as.numeric(as.character(dat$weighting))
dat <- dat[grepl("^[A-Z]{1,8}:US$", dat$bloomberg_symbol), ]

weights <- dat %>% group_by(composite_ticker) %>% summarize(sum = sum(weighting))
dat <- dat %>% group_by(composite_ticker) %>% mutate(total_weight = sum(weighting), norm_weight = weighting/total_weight)
flagged <- unique(dat$composite_ticker[dat$norm_weight >= 1 | dat$norm_weight <= 0])
dat <- dat %>% filter(!composite_ticker %in% flagged)
flagged <- unique(dat$composite_ticker[is.na(dat$norm_weight)])
dat <- dat %>% filter(!composite_ticker %in% flagged)
dat <- dat %>% group_by(composite_ticker) %>% mutate(n = n())

long_mtx <- dat %>% 
  filter(n <= 60) %>% 
  select(composite_ticker, bloomberg_symbol, norm_weight) %>% 
  distinct() %>% 
  unite("ticker", composite_ticker, bloomberg_symbol, remove = FALSE) %>%
  group_by(ticker) %>% 
  top_n(1, norm_weight)

wide <- reshape2::dcast(long_mtx, composite_ticker ~ bloomberg_symbol, value.var = "norm_weight", fill = 0)
tickers <- wide[, 1, drop = T]
wide <- sapply(wide[, -1], as.numeric)

pca <- prcomp(wide, scale = F, rank. = 100)
umap <- uwot::umap(pca$x[, 1:100], n_neighbors = 20, spread = 2, min_dist = 0.2)
umap <- uwot::umap(wide, n_neighbors = 20, spread = 2, min_dist = 0.2)
umap <- as.data.frame(umap)
umap$fund <- tickers
umap$color <- ifelse(grepl("ARKG|GNOM|IDNA|HELX", umap$fund), TRUE, FALSE)
umap <- umap %>% arrange(color)
a <- umap %>% 
  ggplot(aes(x = V1, y = V2, fill = color, tooltip = fund, data_id = fund)) +
  geom_point_interactive(dat = umap %>% filter(color == FALSE), shape = 21, color = "black", size = 2, alpha = 0.6) +
  geom_point_interactive(dat = umap %>% filter(color == TRUE), shape = 21, color = "black", size = 3, alpha = 0.6) +
  scale_fill_manual(values = c("#94A1B2", "#2DB67D")) +
  theme_minimal(18) +
  theme(
    panel.border = element_rect(size = 1, color = "black", fill = "transparent"),
    legend.position = "none",
    axis.title.y = element_text(hjust = 0),
    axis.title.x = element_text(hjust = 0),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.caption = element_text(color = "gray60", hjust = 0, margin = margin(12, 0, 0, 0)),
    plot.subtitle = element_text(color = "gray60")
  ) +
  labs(x = "UMAP1", 
       y = "UMAP2", 
       title = "ETF composition embedding", 
       subtitle = "As of 2021-02-12", 
       caption = "Each point represents an ETF\nHighlighted ETFs: ARKG, IDNA, GNOM, HELX")

girafe(ggobj = a)

cors <- cor(t(wide), method = "spearman")
arkg_cors <- cors[23, ]
names(arkg_cors) <- tickers
sort(arkg_cors, decreasing = TRUE)[1:20]

```

We can see additional funds embedded nearby, such as PTH, PBE, BBH, FBT, EDOC, CHIH, CHNA, and CHB. Additionally, there is a nearby cluster of pharmaceutical ETFs, such as IXJ, PPH, XPH, PJP, IHE, and FTXH. This is an appealing, meaningful visualization of relative ETF holdings. This also agrees with the simpler approach: calculating the Spearman correlation of holdings. Let's focus on our original curated funds and PTH, which I missed in my initial searches.

# Holdings Comparison
Second, we ask what stocks are most popular amongst these funds? For this question, we visualize the number of funds holding a particular equity (point size) across the median holdings weight (y-axis) and market cap (x-axis). This reveals perhaps expected strong biotechnology picks: Illumina, PacBio, Amgen, Bristol-Myers Squibb, Vertex, Regeneron, Gilead, Moderna, Teladoc, Nanostring, CareDx, Fate Therapeutics.

```{r echo=FALSE}

# define funds of interest
thematic_funds <- c("ARKG", "HELX", "GNOM", "PTH", "BBH", "FBT", "CNCR")

sum_tickers <- dat %>%
  filter(composite_ticker %in% thematic_funds) %>%
  group_by(bloomberg_symbol) %>%
  summarize(n = n(), weight = median(norm_weight), mean = log10(mean(market_cap, na.rm = TRUE)))
rmarkdown::paged_table(sum_tickers)

funds_str <- paste0(thematic_funds, collapse = ", ")
sum_tickers$mean[is.nan(sum_tickers$mean)] <- 7
b <- ggplot(data = sum_tickers, 
            mapping = aes(x = mean, y = weight, size = n, tooltip = bloomberg_symbol, data_id = bloomberg_symbol)) +
  geom_point_interactive(shape = 21, color = "black", stroke = 0.75, fill = "gray80", alpha = 0.5) +
  scale_size_continuous(range = c(1, 4)) +
  scale_x_continuous(labels = c("Missing", 8, 9, 10, 11, 12)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "log10(Market Cap $USD)", y = "Median weight",
       title = "Thematic Biotech. ETF Holdings",
       caption = glue::glue("Included funds - {funds_str}")) +
  theme_classic(base_size = 18) +
  ggeasy::easy_all_text_color("black") +
  theme(
    plot.caption = element_text(color = "gray60", size = 8, margin = margin(12, 0, 0, 0)),
    axis.title.x = element_text(margin = margin(12, 0, 0, 0)),
    axis.title.y = element_text(margin = margin(0, 12, 0, 0)),
    panel.background = element_rect(fill = "transparent", color = "black", size = 1),
    plot.background = element_rect(fill = "transparent", color = "transparent", size = 0),
    legend.background = element_rect(fill = "transparent", color = "transparent", size = 0),
    legend.box.background = element_rect(fill = "transparent", color = "transparent", size = 0),
    panel.grid = element_blank(),
    axis.line = element_blank()
  )

girafe(ggobj = b)

```

Lastly, we ask what companies are differentiating ARKG currently? For this question, we simply extract holdings that are unique to ARKG in terms of weight or ownership. Codexis, Adaptive Biotechnologies, Surface Oncology, and Repare Therapeutics are interesting ARKG investments worth a gander.

```{r echo=FALSE}

arkg_picks <- dat %>% 
  filter(composite_ticker %in% thematic_funds) %>% 
  group_by(bloomberg_symbol) %>% 
  mutate(n = n())
arkg_picks <- arkg_picks %>% filter(n == 1 & composite_ticker == "ARKG")
rmarkdown::paged_table(arkg_picks %>% select(bloomberg_symbol, constituent_name, market_cap, norm_weight))

```

We’ve compiled compositional data of ETFs, focused on biotechnology and related themes, and extracted a set of equities selected uniquely or collectively. Much more to dive into from here, but a broad look at ETF holdings in my field is a good start for my foray in the financial and strategic side of biotechnology.

*Disclosures*  
This article expresses my own personal opinions. I am receiving no compensation for it and have no relationship with companies discussed. Nothing discussed is a recommendation to buy, sell, or hold any particular security. I am long ARKG.
